\huge{ \emph{Supplement to} Algorithms For Shaping a Particle Swarm\\ With a Shared Control Input Using Boundary Interaction}

### 2016-12-09 RA-L/experiment.tex
Small changes and the newest version to review
grammar check, force to 8 pages, and revised review doc

\begin{figure*}[!htb]
\centering
\renewcommand{\figwid}{0.38\columnwidth}
{
\begin{overpic}[width =\figwid]{twoR1.pdf}%\put(12,70){$t$  = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR2.pdf}\put(12,70){$t$  = 60 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR3.pdf}\put(12,70){$t$  = 150 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR4.pdf}\put(12,70){$t$  = 160 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR5.pdf}\put(12,70){$t$  = 210 s}
\end{overpic}}
%\vspace{-3em}
\caption{\label{fig:storyReal}Position control of two kilobots (Alg. \ref{alg:XControl}) steered to corresponding colored circle. Boundary walls have nearly infinite friction, so the green robot is stopped by the wall from $t = 60$s until the commanded input is directed away form the wall at $t=150$s, while the pink robot in free-space is unhindered.}
\end{figure*}

@@

Fig. \ref{fig:construction2d} follows the motion of the sliders through iterations  $k$=1, 2, and 7. All robots receive the same control inputs, but boundary interactions break the control symmetry.  Robots reach their respective goal position in a first-in, first-out arrangement beginning with the bottom-left robot from the staging zone occupying the top-right position of the build zone.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{newExperiment.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering $\approx 50$ kilobot robots to desired covariance. The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds. The actual covariance is shown in different trials. Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
\vspace{-1em}
}
\end{figure}

>>>>

\begin{figure*}[!htb]
\centering
\renewcommand{\figwid}{0.38\columnwidth}
{
\begin{overpic}[width =\figwid]{twoR1.pdf}%\put(12,70){$t$  = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR2.pdf}\put(12,70){$t$  = 60 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR3.pdf}\put(12,70){$t$  = 150 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR4.pdf}\put(12,70){$t$  = 160 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR5.pdf}\put(12,70){$t$  = 210 s}
\end{overpic}}
%\vspace{-3em}
\caption{\label{fig:storyReal}Position control of two kilobots (Alg. \ref{alg:PosControl2Robots}) steered to corresponding colored circle. Boundary walls have nearly infinite friction, so the green robot is stopped by the wall from $t = 60$s until the commanded input is directed away form the wall at $t=150$s, while the pink robot in free-space is unhindered.}
\end{figure*}

@@

Fig. \ref{fig:construction2d} follows the motion of the sliders through iterations  $k$=1, 2, and 7. All robots receive the same control inputs, but boundary interactions break the control symmetry.  Robots reach their goal positions in a first-in, first-out arrangement beginning with the bottom-left robot from the staging zone occupying the top-right position of the build zone.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{newExperiment.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering $\approx 50$ kilobot robots to desired covariance. The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds. The actual covariance is shown in different trials. Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
\vspace{-2.5 em}
}
\end{figure}

### 2016-12-09 RA-L/intro.tex
grammar check, force to 8 pages, and revised review doc

\section{Introduction}\label{sec:Intro}
Particle swarms propelled by a global field are common in applied mathematics, biology, and computer graphics. 
If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
\begin{align}
[\dot{x}_i, \dot{y}_i]^\top = \mathbf{u}(t), \qquad i \in [1,n]. \label{eq:swarmDynamics} %[u_x, u_y]^\top
\end{align}
These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\citep{Purcell1977}. 
 In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\citep{Rubenstein2012}.
 Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
Someday large swarms of robots will be remotely guided
to assemble structures in parallel and 
 through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
 For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
They require techniques to shape the swarm that can reliably exploit large populations despite severe underactuation ($2 \ll2n$).

@@

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes, Sec.  \ref{subsec:WallFriction} introduces general models for boundary interaction.  We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in Sec.  \ref{sec:PostionControl2Robots}, and Sec.  \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
Both algorithms require robots with low process noise and require time proportional to the number of robots, so Sec.  \ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm covariance.
Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments with up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.  We end with directions for future research in Sec.  \ref{sec:conclusion}.
%For controlling $Sec. igma_{xy}$, we prove that the swarm position covariance $Sec. igma_{xy}$ is controllable given boundaries with non-zero friction. 
%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
%We conclude by designing controllers that efficiently regulate $Sec. igma_{xy}$.
%This paper
%(1) proves that the swarm position covariance $Sec. igma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $Sec. igma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $Sec. igma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles. 
%\begin{figure}[t]
%\centering
%\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
%}\vspace{-1em}
%\end{figure}
% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

>>>>

\section{Introduction}\label{sec:Intro}
Particle swarms propelled by a global field are common in applied mathematics, biology, and computer graphics. 
If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
\begin{align}
[\dot{x}_i, \dot{y}_i]^\top = \mathbf{u}(t), \qquad i \in [1,n]. \label{eq:swarmDynamics} %[u_x, u_y]^\top
\end{align}
These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\citep{Purcell1977}. 
 In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\citep{Rubenstein2012}.
 Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
Someday large swarms of robots will be remotely guided
to assemble structures in parallel and 
 through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
 For each task, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
They require techniques to shape the swarm that can reliably exploit large populations despite severe underactuation ($2 \ll2n$).

@@

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes, Sec.  \ref{subsec:WallFriction} introduces general models for boundary interaction.  We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in Sec.  \ref{sec:PostionControl2Robots}, and Sec.  \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
Both algorithms require robots with low process noise and require time proportional to the number of robots, so Sec.  \ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm correlation.
Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments with up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.  We end with directions for future research in Sec.  \ref{sec:conclusion}.
%For controlling $Sec. igma_{xy}$, we prove that the swarm position covariance $Sec. igma_{xy}$ is controllable given boundaries with non-zero friction. 
%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
%We conclude by designing controllers that efficiently regulate $Sec. igma_{xy}$.
%This paper
%(1) proves that the swarm position covariance $Sec. igma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $Sec. igma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $Sec. igma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles. 
%\begin{figure}[t]
%\centering
%\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
%}\vspace{-1em}
%\end{figure}
% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

### 2016-12-09 RA-L/simulation.tex
grammar check, force to 8 pages, and revised review doc

Algorithms \ref{alg:PosControl2Robots} %\ref{alg:YControl}, 
was implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm. 
Robot initial positions are shown by crosshairs, and final positions by circled crosshairs.  Dashed lines show the shortest route if robots could be controlled independently.  The path given by  Alg.\ \ref{alg:PosControl2Robots} is shown with solid arrows.
%Each row has five snapshots taken every quarter second. 
For the sake of brevity axis-aligned moves were replaced with oblique moves that combine two moves simultaneously.
% THES IMAGES DON'T DO THIS ANYMORE 
% $\Delta r_x$ is adjusted to $\Delta e_x$ in the second snapshot at $t = 0.25$. 
% The following frames  adjust $\Delta r_y$ to $\Delta e_y$. 
% $\Delta r_y$ is corrected by $t = 0.75$. 
% Finally, the algorithm moves the robots to their corresponding destinations.
An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.

>>>>

Algorithm \ref{alg:PosControl2Robots}  was implemented in Mathematica using point robots (radius = $0$). 
An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.
  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm wiht robot initial positions shown by hollow squares and final positions by circles. 
 Dashed lines show the shortest route if robots could be controlled independently, solid lines the generated path.

   %The path given by our algorithm is shown with solid arrows.
%Each row has five snapshots taken every quarter second. 
%For the sake of brevity axis-aligned moves were replaced with oblique moves that combine 
%jjkjk klklk klklklk klklk lklkl lklklklkl  klklklklk lklklklklk lklklkl klklklkl klklklk klklklk klklk klk klkl
% THES IMAGES DON'T DO THIS ANYMORE 
% $\Delta r_x$ is adjusted to $\Delta e_x$ in the second snapshot at $t = 0.25$. 
% The following frames  adjust $\Delta r_y$ to $\Delta e_y$. 
% $\Delta r_y$ is corrected by $t = 0.75$. 
% Finally, the algorithm moves the robots to their corresponding destinations.

### 2016-12-09 RA-L/theory.tex
grammar check, force to 8 pages, and revised review doc

\subsection{Using Boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary. Like fluid settling in a tank, the stable final configuration minimizes potential energy.
\paragraph{Square workspace}
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$ ranging from [0,1]. Under a global input, they move to a side of the workspace and form a polygonal shape, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.

@@

Computing means, variances, covariance, and correlation requires integrating over the region $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else. 
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:

@@

\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
Global inputs move a swarm uniformly.  
Shape control requires breaking this uniform symmetry.  
A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  
Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
Instead of pushing our robots directly into a wall, this paper examines an oblique approach, by using boundaries that generate friction with the robots.  These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  Robots touching a wall have a friction force that opposes movement along the boundary.  
This causes robots along the boundary to move more slowly than robots in free-space.

@@

The next section shows how a system with friction model \eqref{eq:frictionmodel} and two orthogonal walls can arbitrarily position two robots. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\columnwidth]{friction.pdf} 
\vspace{-1em}
\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
\label{fig:friction}
}
\end{center}
\end{figure}

\section{Algorithms}\label{sec:algorithms}
\subsection{Position Control of $2$ Robots Using Wall Friction}\label{sec:PostionControl2Robots}
\begin{figure*}
\centering
\renewcommand{\figwid}{0.4\columnwidth}
{\begin{overpic}[width =\figwid]{S1_1.png}\put(10,15){$t$ = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_2.png}\put(10,15){$t$ = 0.14 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_3.png}\put(10,15){$t$  = 0.29 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_4.png}\put(10,15){$t$  = 0.43 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_5.png}\put(10,15){$t$  = 0.76 s}
\end{overpic}}\\

@@

Our algorithm solves the positioning problem in four steps: 
First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
Last, the robots, now correctly positioned relative to each other, are moved to their goal locations.

In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all the steps. The worst case path length is $2(\sqrt{2}+1)L$.

\begin{algorithm}
\caption{GenerateDesired spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\scriptsize
\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
 Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$ if not, rotate workspace coordinates $90^{\circ}$.
\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
\State $\epsilon \gets$ robot diameter
\State $\Delta e_y = e_2.y - e_1.y$
\State $\Delta e_x = e_2.x - e_1.x$
%\If {$\Delta e_y \neq 0$}
%\State $sign \gets \frac{|\Delta e_y|}{\Delta e_y}$
%\Else
%\State $sign \gets 0$
%\EndIf
%First Move
\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
%Second Move
%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
%Third Move
\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
%Fourth Move
\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
\If {$\Delta e_y \neq r_2.y - r_1.y$}
%Fifth Move
%TODO This step is not correct.
\State Move $(L- \max(r_1.x, r_2.x) , \epsilon)$ \Comment{Touch right wall}
%Sixth Move
\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ \Comment{Adjust $y$}
\EndIf
%Seventh Move
\State Move by $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
\end{algorithmic}
\end{algorithm}

@@

\State move($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
\State move($f_{kx}-r, 0$)

@@

Assume an obstacle-free, bounded, unit-size square workspace. 
 As shown in Fig.~\ref{fig:SquareFill}, the maximum correlation occurs when the swarm is pushed in the direction $\beta = 3\pi/4$. 
 This correlation as a function of swarm area $A$ is never larger than 1/2, and the maximum correlation decays to 0 as $A$ grows to 1. By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is:
\begin{align} \label{eq:GravityCorrelation}
\rho_{xy} =  \begin{cases}  \frac{1}{2}  , &  0\le A\le \frac{1}{2}  \\
% \frac{3 A (2 (A-2) A+1)}{4 A \left(A^2+3 \sqrt{2-2 A}-6\right)-12 \sqrt{2-2 A}+17}-1
 \frac{3 A (2 (A-2) A+1)}{4 A^3-24 A+\sqrt{2} (12 A-12) \sqrt{1-A}+17}-1
 , & \frac{1}{2} \le A\le 1
\end{cases}
\end{align}
 If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate much larger correlations.
Specifically, we can generate correlations larger than 1/2 by using boundary friction if the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick.

>>>>

\subsection{Using Boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary. Like fluid settling in a tank, the stable final configuration minimizes potential energy.
\paragraph{Square workspace}
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$. Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.

@@

Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else. 
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:

@@

\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
Global inputs move a swarm uniformly.  
Shape control requires breaking this uniform symmetry.  
A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
If the swarm can flow around each other, pushing the swarm into a boundary produces the limited set of configurations presented in Sec.~\ref{subsec:FluidInTank}.
%Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
%An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  
%Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
Instead of pushing our robots directly into a wall, the following sections examine an oblique approach  using boundaries that generate friction with the robots. 
 These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  
Robots touching a wall have a friction force that opposes movement along the boundary.  
This causes robots along the boundary to move more slowly than robots in free-space.

@@

The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\columnwidth]{friction.pdf} 
\vspace{-1em}
\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
\label{fig:friction}
}
\end{center}
\end{figure}

\section{Algorithms}\label{sec:algorithms}
\subsection{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
\begin{figure*}
\centering
\renewcommand{\figwid}{0.4\columnwidth}
{\begin{overpic}[width =\figwid]{S1_1.png}\put(10,15){$t$ = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_2.png}\put(10,15){$t$ = 0.14 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_3.png}\put(10,15){$t$  = 0.29 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_4.png}\put(10,15){$t$  = 0.43 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S1_5.png}\put(10,15){$t$  = 0.76 s}
\end{overpic}}\\

@@

Our algorithm solves the positioning problem in four steps: 
First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.

In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps. The worst case path length is $2(\sqrt{2}+1)L$.

\begin{algorithm}
\caption{GenerateDesiredSpacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\scriptsize
\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
 Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates $90^{\circ}$.
\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
\State $\epsilon \gets$ robot diameter
\State $\Delta e_y = e_2.y - e_1.y$
\State $\Delta e_x = e_2.x - e_1.x$
\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
%Second Move
%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$
\Statex \Comment{Adjust  $y$}
%Third Move
\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ 
\Statex \Comment{Touch bottom wall}
%Fourth Move
\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
\If {$\Delta e_y \neq r_2.y - r_1.y$}
%Fifth Move
%TODO This step is not correct.
\State Move $(L- \max(r_1.x, r_2.x) , \epsilon)$ \Comment{Touch right wall}
%Sixth Move
\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ 
\Statex \Comment{Adjust $y$}
\EndIf
%Seventh Move
\State Move $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
\end{algorithmic}
\end{algorithm}

@@

\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
\State Move ($f_{kx}-r, 0$)

@@

Assume an obstacle-free, bounded, unit-size, square workspace. 
 As shown in Fig.~\ref{fig:SquareFill}, the maximum correlation occurs when the swarm is pushed in the direction $\beta = 3\pi/4$. 
 This correlation as a function of swarm area $A$ is never larger than 1/2, and the maximum correlation decays to 0 as $A$ grows to 1. By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is:
\begin{align} \label{eq:GravityCorrelation}
\rho_{xy} =  \begin{cases}  \frac{1}{2}  , &  0\le A\le \frac{1}{2}  \\
% \frac{3 A (2 (A-2) A+1)}{4 A \left(A^2+3 \sqrt{2-2 A}-6\right)-12 \sqrt{2-2 A}+17}-1
 \frac{3 A (2 (A-2) A+1)}{4 A^3-24 A+\sqrt{2} (12 A-12) \sqrt{1-A}+17}-1
 , & \frac{1}{2} \le A\le 1
\end{cases}
\end{align}
 If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate  larger correlations.
 If the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick we can generate correlations larger than 1/2  using boundary friction.
